Efficient Data Provenance Analysis for Concurrent Programs

### 2013-11-14 paper.tex
committing more in the debugging.stuff file and more text, describing Apache2

\paragraph{Other Applications}
In addition to those applications, we envision that a variety of applications
will be enabled by our simple interface.  Anomaly detection and ``likely
invariant'' techniques similar to AVIO~\cite{avio}, Daikon~\cite{daikon}, and
others are a natural fit for \ctraps.  \ctraps provides the support necessary
for implementing specification checkers for specifications related to
inter-thread communication~\cite{velodrome,oshajava}.  Guided testing
tools~\cite{cuzz,chess} and bug detection tools~\cite{ctrigger} focused on
concurrency could also be implemented as \ctraps applications.  \ctraps also
provides an interesting platform for developing new approaches to
record/replay~\cite{chimera,fdr}.  \ctraps is an enabling platform for
developing these or other analyses with low baseline overheads and without the
need to build an entire concurrency analysis from scratch.

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{l|ll|l|ccc|c}
{\bf App} & {\bf Ver. \#} & {\bf Bug \#} &{\bf Type} & {\bf Eff.?} & {\bf Triv.?} & {\bf Usef.?} & {\bf O.T.?}\\ \hline
mysql  &  4.0.12 &  791     & Atom.  &Yes  &No    & Yes   & No  \\%All values usable
pbzip2 &  0.9.4  &  N/A     & Order  &Yes  &No    & Yes   & Yes \\
httrack& 3.43.9  &  N/A     & Order  &Yes  &No    & Yes   & No  \\%Used w/o origin - no help
jdk1.4 &  1.70   &          & Atom.  &Yes  &No    & Yes   & No  \\%All values usable
apache&  2.0.48 &  21285   & Atom.  &Yes  &No    & Yes   & Maybe\\%All values usable
%apache2&  2.2.9  &  45605   & Atom.  &Yes  &No    & Maybe & ??  \\
\end{tabular}
\caption{\label{tab:bugs}{\bf Concurrency debugging with last writer slices.} We describe each bug and summarize the debugging benefit of last writer slices using the evaluation criteria from ~\cite{badapples}.  }
\normalsize
\end{table}

@@

For {\tt apache}, bad value origin tracking may help.  The bug is an atomicity
violation that leads to a double free.  Two threads manipulate an object's
reference count and deallocate the object when the count hits zero. Zero-checks
and deallocation should be atomic, but the code does not ensure they are.  If
threads interleave their code, multiple threads might see zero reference
counts, which is incorrect.  When that happens, both threads try to free the
object and the second free causes a crash.

@@

>>>>

%\paragraph{Other Applications}
%In addition to those applications, we envision that a variety of applications
%will be enabled by our simple interface.  Anomaly detection and ``likely
%invariant'' techniques similar to AVIO~\cite{avio}, Daikon~\cite{daikon}, and
%others are a natural fit for \ctraps.  \ctraps provides the support necessary
%for implementing specification checkers for specifications related to
%inter-thread communication~\cite{velodrome,oshajava}.  Guided testing
%tools~\cite{cuzz,chess} and bug detection tools~\cite{ctrigger} focused on
%concurrency could also be implemented as \ctraps applications.  \ctraps also
%provides an interesting platform for developing new approaches to
%record/replay~\cite{chimera,fdr}.  \ctraps is an enabling platform for
%developing these or other analyses with low baseline overheads and without the
%need to build an entire concurrency analysis from scratch.

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{l|ll|l|ccc|c}
{\bf App} & {\bf Ver. \#} & {\bf Bug \#} &{\bf Type} & {\bf Eff.?} & {\bf Triv.?} & {\bf Usef.?} & {\bf O.T.?}\\ \hline
jdk1.4 &  1.70   &  N/A     & Atom.  &Yes  &No    &Yes   & No  \\%All values usable
pbzip2 &  0.9.4  &  N/A     & Order  &Yes  &No    &Yes   & Yes \\
httrack& 3.43.9  &  N/A     & Order  &Yes  &No    &Yes   & No  \\%Used w/o origin - no help
transmission & 1.42 & N/A   & Order  &Yes  &No    &Yes   & No  \\
mysql  &  4.0.12 &  791     & Atom.  &Yes  &No    &Yes   & No  \\%All values usable
apache1&  2.0.48 &  21285   & Atom.  &Yes  &No    &Yes   & Some \\%All values usable
apache2&  2.2.9 &  45605    & Atom.  &Yes  &No    &Some  & No \\%All values usable
\end{tabular}
\caption{\label{tab:bugs}{\bf Concurrency debugging with last writer slices.} We describe each bug and summarize the debugging benefit of last writer slices using the evaluation criteria from ~\cite{badapples}.  }
\normalsize
\end{table}

@@

For {\tt apache1}, bad value origin tracking may help.  The bug is an atomicity
violation that leads to a double free.  Two threads manipulate an object's
reference count and deallocate the object when the count hits zero. Zero-checks
and deallocation should be atomic, but the code does not ensure they are.  If
threads interleave their code, multiple threads might see zero reference
counts, which is incorrect.  When that happens, both threads try to free the
object and the second free causes a crash.

@@

\paragraph{apache2}
We used last writer slices to debug an atomicity violation in Apache-2.2.9 and
we found that last writer slices help understand the bug, but not how to fix
it.  The bug involves a group of worker threads that handle web requests and a
network listener thread that passes requests to the workers.  The queue tracks
the number of idle workers.  When a worker becomes idle, it increments the
number of idlers and when the listener passes work to a worker, it decrements
that number.  The counter's increments and decrements are synchronized using
condition variables.  If there are no idlers, the listener waits for a signal.
When a worker becomes idle when no others are, it delivers a signal to the
listener.

The problem is that the listener does not check the condition governed by the
condition variable after being signaled on that condition variable.  As a
result, the listener can incorrectly decrement the idler count twice,
underflowing the unsigned value, which then causes an assertion to fail.  When
the failure occurs, the core dump shows the idler count underflowed, revealing
a primary symptom of the bug, but not providing much help.  We checked the last
writer slice for the idler count and it reported that the listener thread last
wrote the idler count at its decrement.  That information shows the programmer
{\em why} the underflow occurred, which is helpful for understanding the
failure.  However, we conclude that last writer slices are only somewhat 
useful in this case.  Solving this complex bug requires understanding not just
the connection between the underflow and the listener's decrement, but also the
synchronization protocol that coordinates that decrement and other accesses to
the variable.
